---
title: "Table of contents"
output:
  html_document:
    toc: true
    toc_depth: 2
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


```{r, include = FALSE}
# README
# required files:
# Dataset1_Yobs.xlsx
# Model1Baseline_Dataset1_Fit1.rds
```


# Color coding
* <span style="color: blueviolet;">blueviolet</span> for estimated parameters
* <span style="color: deeppink;">deeppink</span> for simulated parameters
* <span style="color: darkturquoise;">darkturquoise</span> for predicted dependent variable
* <span style="color: orange;">orange</span> for observed dependent variable


# Hyperparameters
* $\beta_{0,\mu} = \frac{1}{N} \displaystyle\sum_{n=1}^{N} y^{obs}_{n,t=1}$

* $\beta_{0,\sigma} = 10$

* $\beta_{1,\mu} = 0$

* $\beta_{1,\sigma} = 1$

* $\sigma_\sigma = 1$


# NUTS
* Number of chains: $4$

* Number of iterations per chain: $2000$

* Number of warmup iterations per chain: $1000$

* Number of sampling iterations per chain: $1000$

* Initial values for parameters: random initial values


```{r, include = FALSE}
# preparation ####
# set working directory
setwd("C:/Users/Diiim/Documents/GitHub/BayesianGMMs")

# set decimals to digits instead of scientific
options(scipen = 999)

# load packages
library(readxl)
library(rstan)
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = TRUE)


# simulated data ####
# load observed dependent variable
Y_obs <- data.frame(read_excel("SimulationStudyData/Dataset1_Yobs.xlsx",
                               sheet = "Sheet 1"))

# number of individuals
N <- dim(Y_obs)[1]

# number of time periods
no_periods <- dim(Y_obs)[2]

# time periods
time_periods <- 0:(no_periods-1)
X <- matrix(data = time_periods, nrow = N, ncol = no_periods, byrow = TRUE)

# simulated constant
beta_0_sim <- 10

# simulated linear trend component
beta_1_sim <- 1

# simulated standard deviation for Normal distributions
sigma_sim <- 0.75


# estimated data ####
# load model fit
m_fit <- readRDS("SimulationStudyResults/Model1Baseline_Dataset1_Fit1.rds")
m_fit_params <- rstan::extract(m_fit)

# predicted dependent variable
Y_pred <- m_fit_params$Y_pred

# middle 90% credible intervals for Y_pred
Y_pred_CI_lower <- matrix(data = 0, nrow = N, ncol = no_periods)
Y_pred_CI_upper <- matrix(data = 0, nrow = N, ncol = no_periods)
for (t in 1:no_periods) {
  for (n in 1:N) {
    Y_pred_CI_lower[n,t] <- quantile(Y_pred[,n,t], prob = 0.05)
    Y_pred_CI_upper[n,t] <- quantile(Y_pred[,n,t], prob = 0.95)
  }
}


# colors ####
darkturquoise_transp <- rgb(red = 0,
                            green = 206,
                            blue = 209,
                            max = 209,
                            alpha = 50,
                            names = "darkturquoise")
```


# Convergence
## Rhat, Bulk ESS, and Tail ESS
```{r, include = FALSE}
params <- c("beta_0","beta_1","sigma")
diagnostics <- c("Rhat","Bulk_ESS","Tail_ESS")
Rhat_ESS <-
  as.data.frame(monitor(extract(m_fit, permuted = FALSE)))[params,diagnostics]

```


```{r}
Rhat_ESS
```


## Trace plots
```{r}
# constant
plot(m_fit_params$beta_0,
     type = "l",
     col = "blueviolet",
     main = "",
     xlab = "",
     ylab = "")

grid(nx = NA, ny = NULL)
par(new = TRUE)

plot(m_fit_params$beta_0,
     type = "l",
     col = "blueviolet",
     main = "Constant",
     xlab = "Sampling iteration",
     ylab = "beta_0")

box()
```


```{r}
# linear trend component
plot(m_fit_params$beta_1,
     type = "l",
     col = "blueviolet",
     main = "",
     xlab = "",
     ylab = "")

grid(nx = NA, ny = NULL)
par(new = TRUE)

plot(m_fit_params$beta_1,
     type = "l",
     col = "blueviolet",
     main = "Linear trend component",
     xlab = "Sampling iteration",
     ylab = "beta_1")

box()
```


```{r}
# standard deviation for Y Normal distributions
plot(m_fit_params$sigma,
     type = "l",
     col = "blueviolet",
     main = "",
     xlab = "",
     ylab = "")

grid(nx = NA, ny = NULL)
par(new = TRUE)

plot(m_fit_params$sigma,
     type = "l",
     col = "blueviolet",
     main = "Standard deviation for Y Normal distributions",
     xlab = "Sampling iteration",
     ylab = "sigma")

box()
```


# Posterior
```{r}
# constant
hist(m_fit_params$beta_0,
     xlim = c(min(m_fit_params$beta_0,beta_0_sim),
              max(m_fit_params$beta_0,beta_0_sim)),
     col = "blueviolet",
     border = FALSE,
     main = "",
     xlab = "")

grid()
par(new = TRUE)

hist(m_fit_params$beta_0,
     xlim = c(min(m_fit_params$beta_0,beta_0_sim),
              max(m_fit_params$beta_0,beta_0_sim)),
     col = "blueviolet",
     border = FALSE,
     main = "Constant",
     xlab = "beta_0")

abline(v = beta_0_sim,
       lwd = 2,
       col = "deeppink")

box()
```


```{r}
# linear trend component
hist(m_fit_params$beta_1,
     xlim = c(min(m_fit_params$beta_1,beta_1_sim),
              max(m_fit_params$beta_1,beta_1_sim)),
     col = "blueviolet",
     border = FALSE,
     main = "",
     xlab = "")

grid()
par(new = TRUE)

hist(m_fit_params$beta_1,
     xlim = c(min(m_fit_params$beta_1,beta_1_sim),
              max(m_fit_params$beta_1,beta_1_sim)),
     col = "blueviolet",
     border = FALSE,
     main = "Linear trend component",
     xlab = "beta_1")

abline(v = beta_1_sim,
       lwd = 2,
       col = "deeppink")

box()
```


```{r}
# standard deviation for Y Normal distributions
hist(m_fit_params$sigma,
     xlim = c(min(m_fit_params$sigma,sigma_sim),
              max(m_fit_params$sigma,sigma_sim)),
     col = "blueviolet",
     border = FALSE,
     main = "",
     xlab = "")

grid()
par(new = TRUE)

hist(m_fit_params$sigma,
     xlim = c(min(m_fit_params$sigma,sigma_sim),
              max(m_fit_params$sigma,sigma_sim)),
     col = "blueviolet",
     border = FALSE,
     main = "Standard deviation for Y Normal distributions",
     xlab = "sigma")

abline(v = sigma_sim,
       lwd = 2,
       col = "deeppink")

box()
```


# Posterior predictive check via middle 90% credible interval
```{r}
for (n in 1:N) {
  x <- X[n,]
  y_lower <- Y_pred_CI_lower[n,]
  y_upper <- Y_pred_CI_upper[n,]
  
  plot(x = x,
       y = c(min(y_lower),
             max(y_upper),
             min(Y_obs[n,]),
             max(Y_obs[n,]),
             rep(min(y_lower), times = no_periods-4)),
       type="l",
       col = "white",
       main = paste("n =", n),
       xlab = "t",
       ylab = "y")
  
  grid(nx = NA, ny = NULL)
  par(new = TRUE)
  
  polygon(x = c(x, rev(x)),
          y = c(y_lower, rev(y_upper)),
          col = darkturquoise_transp,
          lty = 0)
  
  lines(x = x,
        y = Y_obs[n,],
        lwd = 2,
        col = "orange")
  
  box()
}
```


