---
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: TRUE
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


```{r, include = FALSE}
# README
# required files from SimulationStudyData/Model1 folder:
# Dataset1_Yobs_Run1.xlsx
# Dataset1_Yobs_Run2.xlsx
# Dataset1_Yobs_Run3.xlsx
# Dataset1_Yobs_Run4.xlsx
# Dataset1_Yobs_Run5.xlsx
# Dataset2_Yobs.xlsx
# Dataset3_Yobs.xlsx
# Dataset4_Yobs.xlsx
# Dataset5_Yobs.xlsx
# Dataset6_Yobs.xlsx

# required files from SimulationStudyData/Model3 folder:
# Dataset15_Yobs.xlsx
# Dataset16_Yobs.xlsx
# Dataset16_Yobslog.xlsx
# Dataset17_Yobs_Run1.xlsx
# Dataset17_Yobs_Run2.xlsx
# Dataset17_Yobs_Run3.xlsx
# Dataset17_Yobs_Run4.xlsx
# Dataset17_Yobs_Run5.xlsx
# Dataset17_Yobslog_Run1.xlsx
# Dataset17_Yobslog_Run2.xlsx
# Dataset17_Yobslog_Run3.xlsx
# Dataset17_Yobslog_Run4.xlsx
# Dataset17_Yobslog_Run5.xlsx
```


```{r, include = FALSE}
# preparation
# set working directory
setwd("C:/Users/Diiim/Documents/Github/BayesianGMMs/SimulationStudyData")

# set decimals to digits instead of scientific
options(scipen = 999)

# load packages
library(readxl)
```


# Model 1
## Dataset 1 - baseline
For $n$ in $1,...,50$,

$\boldsymbol{y}^{obs}_{n} \sim Normal(10 + 1 \, \boldsymbol{x}_{n},0.75)\text{,}$

where $\boldsymbol{x}_{n} = (0,1,2,3,4,5,6,7,8,9)$


```{r}
# number of simulation runs
R <- 5

for (r in 1:R) {
  
  # load observed dependent variable
  Y_obs <- data.frame(read_excel(paste("Model1/Dataset1_Yobs_Run", r, ".xlsx",
                                       sep = ""),
                                 sheet = "Sheet 1"))
  
  # number of individuals
  N <- dim(Y_obs)[1]
  
  # number of time periods
  no_periods <- dim(Y_obs)[2]
  
  # time periods
  time_periods <- 0:(no_periods-1)
  
  # plot Y_obs
  plot(x = time_periods,
       y = c(min(Y_obs),
             max(Y_obs),
             rep(min(Y_obs), times = no_periods-2)),
       type = "l",
       col = "white",
       main = paste("Observed dependent variable for simulation run", r),
       xlab = "x",
       ylab = "y_obs")
  
  grid(nx = NA, ny = NULL)
  par(new = TRUE)
  
  for (n in 1:N) {
    lines(x = time_periods,
          y = Y_obs[n,],
          col = "orange")
  }
  
  box()
  
}
```


## Dataset 2 - two classes - no overlaps between classes
For $n$ in $1,...,200$,

* $\boldsymbol{y}^{obs}_{n} \sim Normal(-5 - 0.5 \, \boldsymbol{x}_{n},0.25) \quad \text{if } z_n = 1$

* $\boldsymbol{y}^{obs}_{n} \sim Normal(10 + 1 \, \boldsymbol{x}_{n},0.75) \quad \text{if } z_n = 2$

* $z_n \sim Categorical(0.3,0.7)\text{,}$

where $\boldsymbol{x}_{n} = (0,1,2,3,4,5,6,7,8,9)$


```{r}
# load observed dependent variable
Y_obs <- data.frame(read_excel("Model1/Dataset2_Yobs.xlsx", sheet = "Sheet 1"))

# number of individuals
N <- dim(Y_obs)[1]

# number of time periods
no_periods <- dim(Y_obs)[2]

# time periods
time_periods <- 0:(no_periods-1)

# plot Y_obs
plot(x = time_periods,
     y = c(min(Y_obs),
           max(Y_obs),
           rep(min(Y_obs), times = no_periods-2)),
     type = "l",
     col = "white",
     main = "Observed dependent variable",
     xlab = "x",
     ylab = "y_obs")
  
grid(nx = NA, ny = NULL)
par(new = TRUE)
  
for (n in 1:N) {
  lines(x = time_periods,
        y = Y_obs[n,],
        col = "orange")
}

box()
```


## Dataset 3 - two classes - overlapping constants
For $n$ in $1,...,200$,

* $\boldsymbol{y}^{obs}_{n} \sim Normal(2 - 0.5 \, \boldsymbol{x}_{n},0.25) \quad \text{if } z_n = 1$

* $\boldsymbol{y}^{obs}_{n} \sim Normal(3 + 1 \, \boldsymbol{x}_{n},0.75) \quad \text{if } z_n = 2$

* $z_n \sim Categorical(0.3,0.7)\text{,}$

where $\boldsymbol{x}_{n} = (0,1,2,3,4,5,6,7,8,9)$


```{r}
# load observed dependent variable
Y_obs <- data.frame(read_excel("Model1/Dataset3_Yobs.xlsx", sheet = "Sheet 1"))

# number of individuals
N <- dim(Y_obs)[1]

# number of time periods
no_periods <- dim(Y_obs)[2]

# time periods
time_periods <- 0:(no_periods-1)

# plot Y_obs
plot(x = time_periods,
     y = c(min(Y_obs),
           max(Y_obs),
           rep(min(Y_obs), times = no_periods-2)),
     type = "l",
     col = "white",
     main = "Observed dependent variable",
     xlab = "x",
     ylab = "y_obs")
  
grid(nx = NA, ny = NULL)
par(new = TRUE)
  
for (n in 1:N) {
  lines(x = time_periods,
        y = Y_obs[n,],
        col = "orange")
}

box()
```


## Dataset 4 - two classes - intersecting trends
For $n$ in $1,...,200$,

* $\boldsymbol{y}^{obs}_{n} \sim Normal(-5 + 1 \, \boldsymbol{x}_{n},0.25) \quad \text{if } z_n = 1$

* $\boldsymbol{y}^{obs}_{n} \sim Normal(10 - 2 \, \boldsymbol{x}_{n},0.75) \quad \text{if } z_n = 2$

* $z_n \sim Categorical(0.3,0.7)\text{,}$

where $\boldsymbol{x}_{n} = (0,1,2,3,4,5,6,7,8,9)$


```{r}
# load observed dependent variable
Y_obs <- data.frame(read_excel("Model1/Dataset4_Yobs.xlsx", sheet = "Sheet 1"))

# number of individuals
N <- dim(Y_obs)[1]

# number of time periods
no_periods <- dim(Y_obs)[2]

# time periods
time_periods <- 0:(no_periods-1)

# plot Y_obs
plot(x = time_periods,
     y = c(min(Y_obs),
           max(Y_obs),
           rep(min(Y_obs), times = no_periods-2)),
     type = "l",
     col = "white",
     main = "Observed dependent variable",
     xlab = "x",
     ylab = "y_obs")
  
grid(nx = NA, ny = NULL)
par(new = TRUE)
  
for (n in 1:N) {
  lines(x = time_periods,
        y = Y_obs[n,],
        col = "orange")
}

box()
```


## Dataset 5 - three classes - no overlaps between classes
For $n$ in $1,...,250$,

* $\boldsymbol{y}^{obs}_{n} \sim Normal(-5 - 0.5 \, \boldsymbol{x}_{n},0.25) \quad \text{if } z_n = 1$

* $\boldsymbol{y}^{obs}_{n} \sim Normal(2.5 + 0.5 \, \boldsymbol{x}_{n},0.5) \quad \text{if } z_n = 2$

* $\boldsymbol{y}^{obs}_{n} \sim Normal(10 + 1 \, \boldsymbol{x}_{n},0.75) \quad \text{if } z_n = 3$

* $z_n \sim Categorical(0.3,0.2,0.5)\text{,}$

where $\boldsymbol{x}_{n} = (0,1,2,3,4,5,6,7,8,9)$


```{r}
# load observed dependent variable
Y_obs <- data.frame(read_excel("Model1/Dataset5_Yobs.xlsx", sheet = "Sheet 1"))

# number of individuals
N <- dim(Y_obs)[1]

# number of time periods
no_periods <- dim(Y_obs)[2]

# time periods
time_periods <- 0:(no_periods-1)

# plot Y_obs
plot(x = time_periods,
     y = c(min(Y_obs),
           max(Y_obs),
           rep(min(Y_obs), times = no_periods-2)),
     type = "l",
     col = "white",
     main = "Observed dependent variable",
     xlab = "x",
     ylab = "y_obs")
  
grid(nx = NA, ny = NULL)
par(new = TRUE)
  
for (n in 1:N) {
  lines(x = time_periods,
        y = Y_obs[n,],
        col = "orange")
}

box()
```


## Dataset 6 - three classes - overlapping constants
For $n$ in $1,...,250$,

* $\boldsymbol{y}^{obs}_{n} \sim Normal(1.5 - 0.5 \, \boldsymbol{x}_{n},0.25) \quad \text{if } z_n = 1$

* $\boldsymbol{y}^{obs}_{n} \sim Normal(2.5 + 0.5 \, \boldsymbol{x}_{n},0.5) \quad \text{if } z_n = 2$

* $\boldsymbol{y}^{obs}_{n} \sim Normal(3.5 + 1 \, \boldsymbol{x}_{n},0.75) \quad \text{if } z_n = 3$

* $z_n \sim Categorical(0.3,0.2,0.5)\text{,}$

where $\boldsymbol{x}_{n} = (0,1,2,3,4,5,6,7,8,9)$


```{r}
# load observed dependent variable
Y_obs <- data.frame(read_excel("Model1/Dataset6_Yobs.xlsx", sheet = "Sheet 1"))

# number of individuals
N <- dim(Y_obs)[1]

# number of time periods
no_periods <- dim(Y_obs)[2]

# time periods
time_periods <- 0:(no_periods-1)

# plot Y_obs
plot(x = time_periods,
     y = c(min(Y_obs),
           max(Y_obs),
           rep(min(Y_obs), times = no_periods-2)),
     type = "l",
     col = "white",
     main = "Observed dependent variable",
     xlab = "x",
     ylab = "y_obs")
  
grid(nx = NA, ny = NULL)
par(new = TRUE)
  
for (n in 1:N) {
  lines(x = time_periods,
        y = Y_obs[n,],
        col = "orange")
}

box()
```


# Model 3
## Dataset 15 - baseline
For $n$ in $1,...,50$,

$\boldsymbol{y}^{obs}_{n} \sim PoissonLog(5 + 0.1 \, \boldsymbol{x}_{n})\text{,}$

where $\boldsymbol{x}_{n} = (0,1,2,3,4,5,6,7,8,9)$


```{r}
# load observed dependent variable
Y_obs <- data.frame(read_excel("Model3/Dataset15_Yobs.xlsx", sheet = "Sheet 1"))

# number of individuals
N <- dim(Y_obs)[1]

# number of time periods
no_periods <- dim(Y_obs)[2]

# time periods
time_periods <- 0:(no_periods-1)

# plot Y_obs
plot(x = time_periods,
     y = c(min(Y_obs),
           max(Y_obs),
           rep(min(Y_obs), times = no_periods-2)),
     type = "l",
     col = "white",
     main = "Observed dependent variable",
     xlab = "x",
     ylab = "y_obs")
  
grid(nx = NA, ny = NULL)
par(new = TRUE)
  
for (n in 1:N) {
  lines(x = time_periods,
        y = Y_obs[n,],
        col = "orange")
}

box()
```


## Dataset 16 - two classes - no overlaps between classes
For $n$ in $1,...,200$,

* $\boldsymbol{y}^{obs}_{n} \sim PoissonLog(1 + 0.4 \, \boldsymbol{x}_{n}) \quad \text{if } z_n = 1$

* $\boldsymbol{y}^{obs}_{n} \sim PoissonLog(5 + 0.1 \, \boldsymbol{x}_{n}) \quad \text{if } z_n = 2$

* $z_n \sim Categorical(0.3,0.7)\text{,}$

where $\boldsymbol{x}_{n} = (0,1,2,3,4,5,6,7,8,9)$


```{r}
# load observed dependent variable
Y_obs <- data.frame(read_excel("Model3/Dataset16_Yobs.xlsx", sheet = "Sheet 1"))
Y_obs_log <- data.frame(read_excel("Model3/Dataset16_Yobslog.xlsx", sheet = "Sheet 1"))

# number of individuals
N <- dim(Y_obs)[1]

# number of time periods
no_periods <- dim(Y_obs)[2]

# time periods
time_periods <- 0:(no_periods-1)

# plot Y_obs
plot(x = time_periods,
     y = c(min(Y_obs),
           max(Y_obs),
           rep(min(Y_obs), times = no_periods-2)),
     type = "l",
     col = "white",
     main = "Observed dependent variable",
     xlab = "x",
     ylab = "y_obs")
  
grid(nx = NA, ny = NULL)
par(new = TRUE)
  
for (n in 1:N) {
  lines(x = time_periods,
        y = Y_obs[n,],
        col = "orange")
}

box()

# plot Y_obs_log
plot(x = time_periods,
     y = c(min(Y_obs_log),
           max(Y_obs_log),
           rep(min(Y_obs_log), times = no_periods-2)),
     type = "l",
     col = "white",
     main = "Observed dependent variable on log scale",
     xlab = "x",
     ylab = "y_obs_log")
  
grid(nx = NA, ny = NULL)
par(new = TRUE)
  
for (n in 1:N) {
  lines(x = time_periods,
        y = Y_obs_log[n,],
        col = "orange")
}

box()
```


## Dataset 17 - two classes - overlapping constants
For $n$ in $1,...,200$,

* $\boldsymbol{y}^{obs}_{n} \sim PoissonLog(1 + 0.4 \, \boldsymbol{x}_{n}) \quad \text{if } z_n = 1$

* $\boldsymbol{y}^{obs}_{n} \sim PoissonLog(1.5 + 0.8 \, \boldsymbol{x}_{n}) \quad \text{if } z_n = 2$

* $z_n \sim Categorical(0.3,0.7)\text{,}$

where $\boldsymbol{x}_{n} = (0,1,2,3,4,5,6,7,8,9)$


```{r}
# number of simulation runs
R <- 5

for (r in 1:R) {
  
  # load observed dependent variable
  Y_obs <- data.frame(read_excel(paste("Model3/Dataset17_Yobs_Run", r, ".xlsx",
                                       sep = ""),
                                 sheet = "Sheet 1"))
  Y_obs <- data.frame(read_excel(paste(
    "Model3/Dataset17_Yobslog_Run", r, ".xlsx",
    sep = ""),
    sheet = "Sheet 1"))
  
  # number of individuals
  N <- dim(Y_obs)[1]
  
  # number of time periods
  no_periods <- dim(Y_obs)[2]
  
  # time periods
  time_periods <- 0:(no_periods-1)
  
  # plot Y_obs
  plot(x = time_periods,
       y = c(min(Y_obs),
             max(Y_obs),
             rep(min(Y_obs), times = no_periods-2)),
       type = "l",
       col = "white",
       main = paste("Observed dependent variable for simulation run", r),
       xlab = "x",
       ylab = "y_obs")
  
  grid(nx = NA, ny = NULL)
  par(new = TRUE)
  
  for (n in 1:N) {
    lines(x = time_periods,
          y = Y_obs[n,],
          col = "orange")
  }
  
  box()
  
  # plot Y_obs_log
  plot(x = time_periods,
       y = c(min(Y_obs_log),
             max(Y_obs_log),
             rep(min(Y_obs_log), times = no_periods-2)),
       type = "l",
       col = "white",
       main = paste(
         "Observed dependent variable on log scale for simulation run", r),
       xlab = "x",
       ylab = "y_obs_log")
  
  grid(nx = NA, ny = NULL)
  par(new = TRUE)
  
  for (n in 1:N) {
    lines(x = time_periods,
          y = Y_obs_log[n,],
          col = "orange")
    }
  
  box()
  
}
```


