---
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: TRUE
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


```{r, include = FALSE}
# README
# required files from SimulationStudyData/Model1 folder:
# SimCase1_Yobs_SimRun1.xlsx
# SimCase2_Yobs_SimRun1.xlsx
# SimCase5_Yobs_SimRun1.xlsx

# required files from SimulationStudyData/Model2 folder:
# SimCase8_Yobs_SimRun1.xlsx
# SimCase9_Yobs_SimRun1.xlsx

# required files from SimulationStudyData/Model3 folder:
# SimCase15_Yobs_SimRun1.xlsx
# SimCase16_Yobs_SimRun1.xlsx
# SimCase16_Yobslog_SimRun1.xlsx
# SimCase17_Yobs_SimRun1.xlsx
# SimCase17_Yobslog_SimRun1.xlsx
# SimCase17v2_Yobs_SimRun1.xlsx
# SimCase17v2_Yobslog_SimRun1.xlsx
# SimCase18_Yobs_SimRun1.xlsx
# SimCase18_Yobslog_SimRun1.xlsx
# SimCase19_Yobs_SimRun1.xlsx
# SimCase19_Yobslog_SimRun1.xlsx

# required files from SimulationStudyData/Model5 folder:
# SimCase29_Yobs_SimRun1.xlsx
# SimCase32_Yobs_SimRun1.xlsx
```


```{r, include = FALSE}
# set working directory
setwd("C:/Users/Diiim/Documents/Github/BayesianGMMs/SimulationStudyData")

# set decimals to digits instead of scientific
options(scipen = 999)

# load packages
library(flextable)
library(readxl)
```


# Overview model 1
```{r}
# simulation case
sim_case <- c("1",
              "2",
              "5")

# number of latent classes
C <- c("baseline",
       "2",
       "3")

# description
descrip <- c("-",
             "classes do not overlap",
             "classes do not overlap")

# number of simulation runs
R <- c(5,
       5,
       5)

# table
df <- data.frame(sim_case,
                 C,
                 descrip,
                 R)

# column names for table
names(df) <- c("Simulation case",
               "Number latent classes",
               "Description",
               "Number simulation runs")

# print table
flextable(df)
```


# Overview model 2
```{r}
# simulation case
sim_case <- c("8",
              "9")

# number of latent classes
C <- c("baseline",
       "2")

# description
descrip <- c("",
             "classes do not overlap")

# number of simulation runs
R <- c(5,
       5)

# table
df <- data.frame(sim_case,
                 C,
                 descrip,
                 R)

# column names for table
names(df) <- c("Simulation case",
               "Number latent classes",
               "Description",
               "Number simulation runs")

# print table
flextable(df)
```


# Overview model 3
```{r}
# simulation case
sim_case <- c("15",
              "16",
              "17",
              "17 v2",
              "18",
              "19")

# number of latent classes
C <- c("baseline",
       "2",
       "2",
       "2",
       "2",
       "3")

# description
descrip <- c("-",
             "classes do not overlap",
             "overlapping constants",
             "overlapping constants",
             "intersecting trends",
             "classes do not overlap")

# number of simulation runs
R <- c(5,
       100,
       100,
       100,
       5,
       5)

# table
df <- data.frame(sim_case,
                 C,
                 descrip,
                 R)

# column names for table
names(df) <- c("Simulation case",
               "Number latent classes",
               "Description",
               "Number simulation runs")

# print table
flextable(df)
```


# Overview model 5
```{r}
# simulation case
sim_case <- c("29",
              "32")

# number of latent classes
C <- c("2",
       "3")

# description
descrip <- c("classes do not overlap",
             "classes do not overlap")

# number of simulation runs
R <- c(5,
       5)

# table
df <- data.frame(sim_case,
                 C,
                 descrip,
                 R)

# column names for table
names(df) <- c("Simulation case",
               "Number latent classes",
               "Description",
               "Number simulation runs")

# print table
flextable(df)
```


# General notation
* Latent class $c$, for $c = 1,...,C$, where $C$ is the number of classes

* Individual $n$, for $n = 1,...,N$, where $N$ is the number of individuals

* Time period $t$, for $t = 1,...,T$, where $T$ is the number of time periods

* $\boldsymbol{Y}^{obs}$ is a $N \times T$ matrix representing an observed dependent variable

* $\boldsymbol{X}$ is a matrix of size $N \times T$ representing an explanatory variable


# Visualizations model 1
## Simulation case 1
Within each simulation run, for $n$ in $1,...,50$ and $t$ in $1,...,10$,

<p style="text-align: center;">$\boldsymbol{y}^{obs}_{n} \sim Normal(10 + 1 \, \boldsymbol{x}_{n},0.75)\text{;}$</p>

where $\boldsymbol{x}_{n} = (0,1,2,3,4,5,6,7,8,9)$.


```{r}
# load observed dependent variable
Y_obs <- data.frame(read_excel("Model1/SimCase1_Yobs_SimRun1.xlsx",
                               sheet = "Sheet 1"))
  
# number of individuals
N <- dim(Y_obs)[1]
  
# number of time periods
no_periods <- dim(Y_obs)[2]
  
# indexes t
indexes_t <- 1:no_periods
  
# plot Y_obs
plot(x = indexes_t,
     y = c(min(Y_obs),
           max(Y_obs),
           rep(min(Y_obs), times = no_periods-2)),
     type = "l",
     col = "white",
     main = "Observed dependent variable for simulation run 1",
     xlab = "t",
     ylab = "y_obs")
  
grid(nx = NA, ny = NULL)
par(new = TRUE)
  
for (n in 1:N) {
  lines(x = indexes_t,
        y = Y_obs[n,],
        col = "orange")
}
  
box()
```


## Simulation case 2
Within each simulation run, for $n$ in $1,...,200$ and $t$ in $1,...,10$,

<p style="text-align: center;">$\boldsymbol{y}^{obs}_{n} \sim Normal(-5 - 0.5 \, \boldsymbol{x}_{n},0.25) \quad \text{if } z^{sim}_n = 1$</p>

<p style="text-align: center;">$\boldsymbol{y}^{obs}_{n} \sim Normal(10 + 1 \, \boldsymbol{x}_{n},0.75) \quad \text{if } z^{sim}_n = 2\text{.}$</p>

$\boldsymbol{z}^{sim}$ is a column vector of size $N$ indicating the simulated class memberships, with

<p style="text-align: center;">$z^{sim}_n \sim Categorical(0.3,0.7)\text{.}$</p>

Furthermore, $\boldsymbol{x}_{n} = (0,1,2,3,4,5,6,7,8,9)$.


```{r}
# load observed dependent variable
Y_obs <- data.frame(read_excel("Model1/SimCase2_Yobs_SimRun1.xlsx",
                               sheet = "Sheet 1"))
  
# number of individuals
N <- dim(Y_obs)[1]
  
# number of time periods
no_periods <- dim(Y_obs)[2]
  
# indexes t
indexes_t <- 1:no_periods
  
# plot Y_obs
plot(x = indexes_t,
     y = c(min(Y_obs),
           max(Y_obs),
           rep(min(Y_obs), times = no_periods-2)),
     type = "l",
     col = "white",
     main = "Observed dependent variable for simulation run 1",
     xlab = "t",
     ylab = "y_obs")
  
grid(nx = NA, ny = NULL)
par(new = TRUE)
  
for (n in 1:N) {
  lines(x = indexes_t,
        y = Y_obs[n,],
        col = "orange")
}
  
box()
```


## Simulation case 5
Within each simulation run, for $n$ in $1,...,250$ and $t$ in $1,...,10$,

<p style="text-align: center;">$\boldsymbol{y}^{obs}_{n} \sim Normal(-5 - 0.5 \, \boldsymbol{x}_{n},0.25) \quad \text{if } z^{sim}_n = 1$</p>

<p style="text-align: center;">$\boldsymbol{y}^{obs}_{n} \sim Normal(2.5 + 0.5 \, \boldsymbol{x}_{n},0.5) \quad \text{if } z^{sim}_n = 2$</p>

<p style="text-align: center;">$\boldsymbol{y}^{obs}_{n} \sim Normal(10 + 1 \, \boldsymbol{x}_{n},0.75) \quad \text{if } z^{sim}_n = 3\text{.}$</p>

$\boldsymbol{z}^{sim}$ is a column vector of size $N$ indicating the simulated class memberships, with

<p style="text-align: center;">$z^{sim}_n \sim Categorical(0.3,0.2,0.5)\text{.}$</p>

Furthermore, $\boldsymbol{x}_{n} = (0,1,2,3,4,5,6,7,8,9)$.


```{r}
# load observed dependent variable
Y_obs <- data.frame(read_excel("Model1/SimCase5_Yobs_SimRun1.xlsx",
                               sheet = "Sheet 1"))
  
# number of individuals
N <- dim(Y_obs)[1]
  
# number of time periods
no_periods <- dim(Y_obs)[2]
  
# indexes t
indexes_t <- 1:no_periods
  
# plot Y_obs
plot(x = indexes_t,
     y = c(min(Y_obs),
           max(Y_obs),
           rep(min(Y_obs), times = no_periods-2)),
     type = "l",
     col = "white",
     main = "Observed dependent variable for simulation run 1",
     xlab = "t",
     ylab = "y_obs")
  
grid(nx = NA, ny = NULL)
par(new = TRUE)
  
for (n in 1:N) {
  lines(x = indexes_t,
        y = Y_obs[n,],
        col = "orange")
}
  
box()
```


# Visualizations model 2
## Simulation case 8
Within each simulation run, for $n$ in $1,...,50$ and $t$ in $1,...,10$,

<p style="text-align: center;">$\boldsymbol{y}^{obs}_{n} \sim Normal(10 + 1 \, \boldsymbol{x}_{n} - 0.1 \, \boldsymbol{x}^2_{n},0.75)\text{;}$</p>

where $\boldsymbol{x}_{n} = (0,1,2,3,4,5,6,7,8,9)$.


```{r}
# load observed dependent variable
Y_obs <- data.frame(read_excel("Model2/SimCase8_Yobs_SimRun1.xlsx",
                               sheet = "Sheet 1"))
  
# number of individuals
N <- dim(Y_obs)[1]
  
# number of time periods
no_periods <- dim(Y_obs)[2]
  
# indexes t
indexes_t <- 1:no_periods
  
# plot Y_obs
plot(x = indexes_t,
     y = c(min(Y_obs),
           max(Y_obs),
           rep(min(Y_obs), times = no_periods-2)),
     type = "l",
     col = "white",
     main = "Observed dependent variable for simulation run 1",
     xlab = "t",
     ylab = "y_obs")
  
grid(nx = NA, ny = NULL)
par(new = TRUE)
  
for (n in 1:N) {
  lines(x = indexes_t,
        y = Y_obs[n,],
        col = "orange")
}
  
box()
```


## Simulation case 9
Within each simulation run, for $n$ in $1,...,200$ and $t$ in $1,...,10$,

<p style="text-align: center;">$\boldsymbol{y}^{obs}_{n} \sim Normal(-5 - 0.5 \, \boldsymbol{x}_{n} + 0.2 \, \boldsymbol{x}^2_{n},0.25) \quad \text{if } z^{sim}_n = 1$</p>

<p style="text-align: center;">$\boldsymbol{y}^{obs}_{n} \sim Normal(10 + 1 \, \boldsymbol{x}_{n} - 0.1 \, \boldsymbol{x}^2_{n},0.75) \quad \text{if } z^{sim}_n = 2\text{.}$</p>

$\boldsymbol{z}^{sim}$ is a column vector of size $N$ indicating the simulated class memberships, with

<p style="text-align: center;">$z^{sim}_n \sim Categorical(0.3,0.7)\text{.}$</p>

Furthermore, $\boldsymbol{x}_{n} = (0,1,2,3,4,5,6,7,8,9)$.


```{r}
# load observed dependent variable
Y_obs <- data.frame(read_excel("Model2/SimCase9_Yobs_SimRun1.xlsx",
                               sheet = "Sheet 1"))
  
# number of individuals
N <- dim(Y_obs)[1]
  
# number of time periods
no_periods <- dim(Y_obs)[2]
  
# indexes t
indexes_t <- 1:no_periods
  
# plot Y_obs
plot(x = indexes_t,
     y = c(min(Y_obs),
           max(Y_obs),
           rep(min(Y_obs), times = no_periods-2)),
     type = "l",
     col = "white",
     main = "Observed dependent variable for simulation run 1",
     xlab = "t",
     ylab = "y_obs")
  
grid(nx = NA, ny = NULL)
par(new = TRUE)
  
for (n in 1:N) {
  lines(x = indexes_t,
        y = Y_obs[n,],
        col = "orange")
}
  
box()
```


# Visualizations model 3
## Simulation case 15
Within each simulation run, for $n$ in $1,...,50$ and $t$ in $1,...,10$,

<p style="text-align: center;">$\boldsymbol{y}^{obs}_{n} \sim PoissonLog(5 + 0.1 \, \boldsymbol{x}_{n})\text{;}$</p>

where $\boldsymbol{x}_{n} = (0,1,2,3,4,5,6,7,8,9)$.


```{r}
# load observed dependent variable
Y_obs <- data.frame(read_excel("Model3/SimCase15_Yobs_SimRun1.xlsx",
                               sheet = "Sheet 1"))
  
# number of individuals
N <- dim(Y_obs)[1]
  
# number of time periods
no_periods <- dim(Y_obs)[2]
  
# indexes t
indexes_t <- 1:no_periods
  
# plot Y_obs
plot(x = indexes_t,
     y = c(min(Y_obs),
           max(Y_obs),
           rep(min(Y_obs), times = no_periods-2)),
     type = "l",
     col = "white",
     main = "Observed dependent variable for simulation run 1",
     xlab = "t",
     ylab = "y_obs")
  
grid(nx = NA, ny = NULL)
par(new = TRUE)
  
for (n in 1:N) {
  lines(x = indexes_t,
        y = Y_obs[n,],
        col = "orange")
}
  
box()
```


## Simulation case 16
Within each simulation run, for $n$ in $1,...,200$ and $t$ in $1,...,10$,

<p style="text-align: center;">$\boldsymbol{y}^{obs}_{n} \sim PoissonLog(1 + 0.4 \, \boldsymbol{x}_{n}) \quad \text{if } z^{sim}_n = 1$</p>

<p style="text-align: center;">$\boldsymbol{y}^{obs}_{n} \sim PoissonLog(5 + 0.1 \, \boldsymbol{x}_{n}) \quad \text{if } z^{sim}_n = 2\text{.}$</p>

$\boldsymbol{z}^{sim}$ is a column vector of size $N$ indicating the simulated class memberships, with

<p style="text-align: center;">$z^{sim}_n \sim Categorical(0.3,0.7)\text{.}$</p>

Furthermore, $\boldsymbol{x}_{n} = (0,1,2,3,4,5,6,7,8,9)$.


```{r}
# load observed dependent variable
Y_obs <- data.frame(read_excel("Model3/SimCase16_Yobs_SimRun1.xlsx",
                               sheet = "Sheet 1"))
Y_obs_log <- data.frame(read_excel("Model3/SimCase16_Yobslog_SimRun1.xlsx",
                                   sheet = "Sheet 1"))
  
# number of individuals
N <- dim(Y_obs)[1]
  
# number of time periods
no_periods <- dim(Y_obs)[2]
  
# indexes t
indexes_t <- 1:no_periods
  
# plot Y_obs
plot(x = indexes_t,
     y = c(min(Y_obs),
           max(Y_obs),
           rep(min(Y_obs), times = no_periods-2)),
     type = "l",
     col = "white",
     main = "Observed dependent variable for simulation run 1",
     xlab = "t",
     ylab = "y_obs")
  
grid(nx = NA, ny = NULL)
par(new = TRUE)
  
for (n in 1:N) {
  lines(x = indexes_t,
        y = Y_obs[n,],
        col = "orange")
}
  
box()
  
# plot Y_obs_log
plot(x = indexes_t,
     y = c(min(Y_obs_log),
           max(Y_obs_log),
           rep(min(Y_obs_log), times = no_periods-2)),
     type = "l",
     col = "white",
     main = "Observed dependent variable on log scale for simulation run 1",
     xlab = "t",
     ylab = "y_obs_log")

grid(nx = NA, ny = NULL)
par(new = TRUE)
  
for (n in 1:N) {
  lines(x = indexes_t,
        y = Y_obs_log[n,],
        col = "orange")
}
  
box()
```


## Simulation case 17
Within each simulation run, for $n$ in $1,...,200$ and $t$ in $1,...,10$,

<p style="text-align: center;">$\boldsymbol{y}^{obs}_{n} \sim PoissonLog(4.75 + 0.05 \, \boldsymbol{x}_{n}) \quad \text{if } z^{sim}_n = 1$</p>

<p style="text-align: center;">$\boldsymbol{y}^{obs}_{n} \sim PoissonLog(5 + 0.1 \, \boldsymbol{x}_{n}) \quad \text{if } z^{sim}_n = 2\text{.}$</p>

$\boldsymbol{z}^{sim}$ is a column vector of size $N$ indicating the simulated class memberships, with

<p style="text-align: center;">$z^{sim}_n \sim Categorical(0.3,0.7)\text{.}$</p>

Furthermore, $\boldsymbol{x}_{n} = (0,1,2,3,4,5,6,7,8,9)$.


```{r}
# load observed dependent variable
Y_obs <- data.frame(read_excel("Model3/SimCase17_Yobs_SimRun1.xlsx",
                               sheet = "Sheet 1"))
Y_obs_log <- data.frame(read_excel("Model3/SimCase17_Yobslog_SimRun1.xlsx",
                                   sheet = "Sheet 1"))
  
# number of individuals
N <- dim(Y_obs)[1]
  
# number of time periods
no_periods <- dim(Y_obs)[2]
  
# indexes t
indexes_t <- 1:no_periods
  
# plot Y_obs
plot(x = indexes_t,
     y = c(min(Y_obs),
           max(Y_obs),
           rep(min(Y_obs), times = no_periods-2)),
     type = "l",
     col = "white",
     main = "Observed dependent variable for simulation run 1",
     xlab = "t",
     ylab = "y_obs")
  
grid(nx = NA, ny = NULL)
par(new = TRUE)
  
for (n in 1:N) {
  lines(x = indexes_t,
        y = Y_obs[n,],
        col = "orange")
}
  
box()
  
# plot Y_obs_log
plot(x = indexes_t,
     y = c(min(Y_obs_log),
           max(Y_obs_log),
           rep(min(Y_obs_log), times = no_periods-2)),
     type = "l",
     col = "white",
     main = "Observed dependent variable on log scale for simulation run 1",
     xlab = "t",
     ylab = "y_obs_log")

grid(nx = NA, ny = NULL)
par(new = TRUE)
  
for (n in 1:N) {
  lines(x = indexes_t,
        y = Y_obs_log[n,],
        col = "orange")
}
  
box()
```


## Simulation case 17 version 2
Within each simulation run, for $n$ in $1,...,200$ and $t$ in $1,...,10$,

<p style="text-align: center;">$\boldsymbol{y}^{obs}_{n} \sim PoissonLog(5 + 0.05 \, \boldsymbol{x}_{n}) \quad \text{if } z^{sim}_n = 1$</p>

<p style="text-align: center;">$\boldsymbol{y}^{obs}_{n} \sim PoissonLog(4.75 + 0.075 \, \boldsymbol{x}_{n}) \quad \text{if } z^{sim}_n = 2\text{.}$</p>

$\boldsymbol{z}^{sim}$ is a column vector of size $N$ indicating the simulated class memberships, with

<p style="text-align: center;">$z^{sim}_n \sim Categorical(0.3,0.7)\text{.}$</p>

Furthermore, $\boldsymbol{x}_{n} = (0,1,2,3,4,5,6,7,8,9)$.


```{r}
# load observed dependent variable
Y_obs <- data.frame(read_excel("Model3/SimCase17v2_Yobs_SimRun1.xlsx",
                               sheet = "Sheet 1"))
Y_obs_log <- data.frame(read_excel("Model3/SimCase17v2_Yobslog_SimRun1.xlsx",
                                   sheet = "Sheet 1"))
  
# number of individuals
N <- dim(Y_obs)[1]
  
# number of time periods
no_periods <- dim(Y_obs)[2]
  
# indexes t
indexes_t <- 1:no_periods
  
# plot Y_obs
plot(x = indexes_t,
     y = c(min(Y_obs),
           max(Y_obs),
           rep(min(Y_obs), times = no_periods-2)),
     type = "l",
     col = "white",
     main = "Observed dependent variable for simulation run 1",
     xlab = "t",
     ylab = "y_obs")
  
grid(nx = NA, ny = NULL)
par(new = TRUE)
  
for (n in 1:N) {
  lines(x = indexes_t,
        y = Y_obs[n,],
        col = "orange")
}
  
box()
  
# plot Y_obs_log
plot(x = indexes_t,
     y = c(min(Y_obs_log),
           max(Y_obs_log),
           rep(min(Y_obs_log), times = no_periods-2)),
     type = "l",
     col = "white",
     main = "Observed dependent variable on log scale for simulation run 1",
     xlab = "t",
     ylab = "y_obs_log")

grid(nx = NA, ny = NULL)
par(new = TRUE)
  
for (n in 1:N) {
  lines(x = indexes_t,
        y = Y_obs_log[n,],
        col = "orange")
}
  
box()
```


## Simulation case 18
Within each simulation run, for $n$ in $1,...,200$ and $t$ in $1,...,10$,

<p style="text-align: center;">$\boldsymbol{y}^{obs}_{n} \sim PoissonLog(1 + 0.4 \, \boldsymbol{x}_{n}) \quad \text{if } z^{sim}_n = 1$</p>

<p style="text-align: center;">$\boldsymbol{y}^{obs}_{n} \sim PoissonLog(5 - 0.2 \, \boldsymbol{x}_{n}) \quad \text{if } z^{sim}_n = 2\text{.}$</p>

$\boldsymbol{z}^{sim}$ is a column vector of size $N$ indicating the simulated class memberships, with

<p style="text-align: center;">$z^{sim}_n \sim Categorical(0.3,0.7)\text{.}$</p>

Furthermore, $\boldsymbol{x}_{n} = (0,1,2,3,4,5,6,7,8,9)$.


```{r}
# load observed dependent variable
Y_obs <- data.frame(read_excel("Model3/SimCase18_Yobs_SimRun1.xlsx",
                               sheet = "Sheet 1"))
Y_obs_log <- data.frame(read_excel("Model3/SimCase18_Yobslog_SimRun1.xlsx",
                                   sheet = "Sheet 1"))
  
# number of individuals
N <- dim(Y_obs)[1]
  
# number of time periods
no_periods <- dim(Y_obs)[2]
  
# indexes t
indexes_t <- 1:no_periods
  
# plot Y_obs
plot(x = indexes_t,
     y = c(min(Y_obs),
             max(Y_obs),
             rep(min(Y_obs), times = no_periods-2)),
     type = "l",
     col = "white",
     main = "Observed dependent variable for simulation run 1",
     xlab = "t",
     ylab = "y_obs")
  
grid(nx = NA, ny = NULL)
par(new = TRUE)
  
for (n in 1:N) {
  lines(x = indexes_t,
        y = Y_obs[n,],
        col = "orange")
}
  
box()
  
# plot Y_obs_log
plot(x = indexes_t,
     y = c(min(Y_obs_log),
           max(Y_obs_log),
           rep(min(Y_obs_log), times = no_periods-2)),
     type = "l",
     col = "white",
     main = "Observed dependent variable on log scale for simulation run 1",
     xlab = "t",
     ylab = "y_obs_log")

grid(nx = NA, ny = NULL)
par(new = TRUE)

for (n in 1:N) {
  lines(x = indexes_t,
        y = Y_obs_log[n,],
        col = "orange")
}
  
box()
```


## Simulation case 19
Within each simulation run, for $n$ in $1,...,250$ and $t$ in $1,...,10$,

<p style="text-align: center;">$\boldsymbol{y}^{obs}_{n} \sim PoissonLog(1 + 0.4 \, \boldsymbol{x}_{n}) \quad \text{if } z^{sim}_n = 1$</p>

<p style="text-align: center;">$\boldsymbol{y}^{obs}_{n} \sim PoissonLog(3 + 0.25 \, \boldsymbol{x}_{n}) \quad \text{if } z^{sim}_n = 2$</p>

<p style="text-align: center;">$\boldsymbol{y}^{obs}_{n} \sim PoissonLog(5 + 0.1 \, \boldsymbol{x}_{n}) \quad \text{if } z^{sim}_n = 3\text{.}$</p>

$\boldsymbol{z}^{sim}$ is a column vector of size $N$ indicating the simulated class memberships, with

<p style="text-align: center;">$z^{sim}_n \sim Categorical(0.3,0.2,0.5)\text{.}$</p>

Furthermore, $\boldsymbol{x}_{n} = (0,1,2,3,4,5,6,7,8,9)$.


```{r}
# load observed dependent variable
Y_obs <- data.frame(read_excel("Model3/SimCase19_Yobs_SimRun1.xlsx",
                               sheet = "Sheet 1"))
Y_obs_log <- data.frame(read_excel("Model3/SimCase19_Yobslog_SimRun1.xlsx",
                                   sheet = "Sheet 1"))
  
# number of individuals
N <- dim(Y_obs)[1]
  
# number of time periods
no_periods <- dim(Y_obs)[2]
  
# indexes t
indexes_t <- 1:no_periods
  
# plot Y_obs
plot(x = indexes_t,
     y = c(min(Y_obs),
           max(Y_obs),
           rep(min(Y_obs), times = no_periods-2)),
     type = "l",
     col = "white",
     main = "Observed dependent variable for simulation run 1",
     xlab = "t",
     ylab = "y_obs")
  
grid(nx = NA, ny = NULL)
par(new = TRUE)
  
for (n in 1:N) {
  lines(x = indexes_t,
        y = Y_obs[n,],
        col = "orange")
}
  
box()
  
# plot Y_obs_log
plot(x = indexes_t,
     y = c(min(Y_obs_log),
           max(Y_obs_log),
           rep(min(Y_obs_log), times = no_periods-2)),
     type = "l",
     col = "white",
     main = "Observed dependent variable on log scale for simulation run 1",
     xlab = "t",
     ylab = "y_obs_log")

grid(nx = NA, ny = NULL)
par(new = TRUE)
  
for (n in 1:N) {
  lines(x = indexes_t,
        y = Y_obs_log[n,],
        col = "orange")
}
  
box()
```


# Visualizations model 5
## Simulation case 29
Within each simulation run, for $n$ in $1,...,200$ and $t$ in $1,...,10$,

<p style="text-align: center;">$\boldsymbol{y}^{obs}_{n,t} \sim Normal(-5 - 0.5 \, \boldsymbol{x}_{n},0.25) \quad \text{if } z^{sim}_{n,t} = 1$</p>

<p style="text-align: center;">$\boldsymbol{y}^{obs}_{n,t} \sim Normal(10 + 1 \, \boldsymbol{x}_{n},0.75) \quad \text{if } z^{sim}_{n,t} = 2\text{.}$</p>

$\boldsymbol{Z}^{sim}$ is a $N \times T$ matrix indicating the simulated class memberships, with

<p style="text-align: center;">$z^{sim}_{n,t} \sim Categorical(0.3,0.7) \quad \text{for } t = 1$</p>

<p style="text-align: center;">$z^{sim}_{n,t} \sim Categorical(\boldsymbol{\psi}^{sim}_{z_{n,t-1}}) \quad \text{for } t > 1\text{.}$</p>

$\boldsymbol{\Psi}^{sim}$ is a simulated $C \times C$ transition matrix, with $\boldsymbol{\psi}^{sim}_1 = (0.99,0.01)$ and $\boldsymbol{\psi}^{sim}_2 = (0.05,0.95)$. Furthermore, $\boldsymbol{x}_{n} = (0,1,2,3,4,5,6,7,8,9)$.


```{r}
# load observed dependent variable
Y_obs <- data.frame(read_excel("Model5/SimCase29_Yobs_SimRun1.xlsx",
                               sheet = "Sheet 1"))
  
# number of individuals
N <- dim(Y_obs)[1]
  
# number of time periods
no_periods <- dim(Y_obs)[2]
  
# indexes t
indexes_t <- 1:no_periods
  
# plot Y_obs
plot(x = indexes_t,
     y = c(min(Y_obs),
           max(Y_obs),
           rep(min(Y_obs), times = no_periods-2)),
     type = "l",
     col = "white",
     main = "Observed dependent variable for simulation run 1",
     xlab = "t",
     ylab = "y_obs")
  
grid(nx = NA, ny = NULL)
par(new = TRUE)
  
for (n in 1:N) {
  lines(x = indexes_t,
        y = Y_obs[n,],
        col = "orange")
}
  
box()
```


## Simulation case 32
Within each simulation run, for $n$ in $1,...,250$ and $t$ in $1,...,10$,

<p style="text-align: center;">$\boldsymbol{y}^{obs}_{n,t} \sim Normal(-5 - 0.5 \, \boldsymbol{x}_{n},0.25) \quad \text{if } z^{sim}_{n,t} = 1$</p>

<p style="text-align: center;">$\boldsymbol{y}^{obs}_{n,t} \sim Normal(2.5 + 0.5 \, \boldsymbol{x}_{n},0.5) \quad \text{if } z^{sim}_{n,t} = 2$</p>

<p style="text-align: center;">$\boldsymbol{y}^{obs}_{n,t} \sim Normal(10 + 1 \, \boldsymbol{x}_{n},0.75) \quad \text{if } z^{sim}_{n,t} = 3\text{.}$</p>

$\boldsymbol{Z}^{sim}$ is a $N \times T$ matrix indicating the simulated class memberships, with

<p style="text-align: center;">$z^{sim}_{n,t} \sim Categorical(0.3,0.2,0.5) \quad \text{for } t = 1$</p>

<p style="text-align: center;">$z^{sim}_{n,t} \sim Categorical(\boldsymbol{\psi}^{sim}_{z_{n,t-1}}) \quad \text{for } t > 1\text{.}$</p>

$\boldsymbol{\Psi}^{sim}$ is a simulated $C \times C$ transition matrix, with $\boldsymbol{\psi}^{sim}_1 = (0.96,0.01,0.03)$, $\boldsymbol{\psi}^{sim}_2 = (0.01,0.97,0.02)$, and $\boldsymbol{\psi}^{sim}_3 = (0.02,0.03,0.95)$. Furthermore, $\boldsymbol{x}_{n} = (0,1,2,3,4,5,6,7,8,9)$.


```{r}
# load observed dependent variable
Y_obs <- data.frame(read_excel("Model5/SimCase32_Yobs_SimRun1.xlsx",
                               sheet = "Sheet 1"))
  
# number of individuals
N <- dim(Y_obs)[1]
  
# number of time periods
no_periods <- dim(Y_obs)[2]
  
# indexes t
indexes_t <- 1:no_periods
  
# plot Y_obs
plot(x = indexes_t,
     y = c(min(Y_obs),
           max(Y_obs),
           rep(min(Y_obs), times = no_periods-2)),
     type = "l",
     col = "white",
     main = "Observed dependent variable for simulation run 1",
     xlab = "t",
     ylab = "y_obs")
  
grid(nx = NA, ny = NULL)
par(new = TRUE)
  
for (n in 1:N) {
  lines(x = indexes_t,
        y = Y_obs[n,],
        col = "orange")
}
  
box()
```


